#case study 1

Link to rshiny app!  https://tinapai.shinyapps.io/rshiny/

```{r}
library(tidyverse)
library(VIM)
library(lattice)
library(DMwR)
library(mice)
library(ggplot2)
library(ggthemes)
library(stringr)
library(class)
library(caret)
```

```{r}
# Reading in the data and observing the first 6 observations for each data set
beers = read.csv("Beers.csv")
breweries = read.csv("Breweries.csv")
head(beers)
head(breweries)
```


#question 1: how many breweries are present in each state?

There are ranging from 1 to 47 breweries in the states of the US. The state with the most breweries is CO (with 47 breweries) and the states with the fewest are NV, DC, ND, and SD (with 1 brewery).

```{r}
# We create a new data set where we group it by State and create a tally of of how many breweries are in each state, arrange it in descending order and show the 6  highest and lowest state frequencies
state_breweries <- breweries %>%
  group_by(State) %>%
  tally() %>%
  arrange(desc(n))
head(state_breweries)
tail(state_breweries)

#We create a bar graph of our previous results

state_breweries %>%
  ggplot(aes(x=reorder(State, -n), y=n)) +
  geom_bar(stat = "identity") +
  ggtitle("Number of Breweries by State") +
  xlab("State") +
  ylab("Number of Breweries")
```

#question 2: merge data

```{r}
#Merging the data sets with key Brew_ID and Brewery_id since they are the same but have different names
head(breweries)
head(beers)
beer_breweries = breweries %>%
  left_join(beers, by = c("Brew_ID" = "Brewery_id"))
names(beer_breweries) <- c("Brew_ID","Brewery_Name","City","State","Beer_Name", "Beer_ID","ABV","IBU","Style","Ounces")
head(beer_breweries)
```


#question 3: fix nas
```{r}

#Plot the NAs and look the proportion of NAs per column
beer_breweries_miss = aggr(beer_breweries, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, labels=names(beer_breweries), cex.axis=.7, gap=3, ylab=c("Proportion of NAs","NA Pattern"))

#Using Knn imputation as means to fill in the NA values
beer_breweries_impute <- knnImputation(beer_breweries)
beer_breweries <- beer_breweries_impute
```


#question 4: Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

Median alcohol content by state ranged from .04 ABV in UT to just over .06 ABV in DC. Below shows graphs of the median ABV by state.

Median bitterness by state ranged from about 20 IBU in KS to just under 60 IBU in WV. Below shows graphs of the median IBU by state.
```{r}
#median abv by state
state_abv = beer_breweries %>%
  group_by(State) %>%
  summarise(median = median(ABV, na.rm = TRUE))
head(state_abv)

#Bar plot of median abv by state
state_abv %>%
  ggplot(aes(x=reorder(State, -median), y=median)) +
  geom_bar(stat = "identity") +
  ggtitle("Median ABV by State") +
  xlab("State") +
  ylab("Median ABV")

#median ibu by state
state_ibu = beer_breweries %>%
  group_by(State) %>%
  summarise(median = median(IBU, na.rm = TRUE))
head(state_ibu)
#Bar plot of median IBU by state
state_ibu %>%
  ggplot(aes(x=reorder(State, -median), y=median)) +
  geom_bar(stat = "identity") +
  ggtitle("Median IBU by State") +
  xlab("State") +
  ylab("Median IBU")
```



#question 5: Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

The beer with most alcoholic content is the Lee Hill Quadruple Ale from Upslope Brewing Company in Colorado, at .128 ABV.

The beer with the most bitterness is Bitter Bitch Imperial IPA from Astoria Brewing Company in Oregon, at 138 IBU.

```{r}
#Create a dataset where we arrange the beer with highest ABV in descending order
#CO has the beer with the highest ABV
top_ABV <- beer_breweries %>% 
  arrange(desc(ABV)) %>%
  top_n(10, ABV)
head(top_ABV[c("State", "Brewery_Name", "Beer_Name", "ABV")])

#Create a dataset where we arrange the beer with highest IBU in descending order
#OR has the beer with the highest IBU
top_IBU <- beer_breweries %>%
  arrange(desc(IBU)) %>%
  top_n(10, IBU)
head(top_IBU[c("State", "Brewery_Name", "Beer_Name", "IBU")])

```


#question 6: Comment on the summary statistics and distribution of the ABV variable.
The ABV ranges from .027 to .125, with a median of .057 and mean of .0599. The distribution looks somewhat right skewed.

```{r}
#Plot a histogram of the ABV distribution
summary(beer_breweries$ABV)
beer_breweries %>%
  ggplot(aes(x=ABV)) +
  geom_histogram() +
  ggtitle("Histogram of ABV")
```


#question 7: Is there a relationship between the bitterness of the beer and its alcoholic content?

There seems to be a positive correlation between ABV and IBU. So that means there is a positive relationship betwen ABV and IBU so as ABV increases, so does IBU.

```{r}
#Plot a scatterplot comparing the two variables ABV and IBU with a regression line
beer_breweries %>% ggplot(aes(x=ABV,y=IBU)) + theme_fivethirtyeight() + geom_point()+geom_smooth(method=lm) +
  xlab("Alcohol By Volume (ABV)") +
  ylab("International Bitterness Unit (IBU)") + 
  ggtitle( "Relationship Between ABV and IBU")

#Calculate the correlation betwen ABV and IBU
cor(beer_breweries$ABV,beer_breweries$IBU)

```

#question 8: IPA vs Ale

```{r}
# We use Regex to find which style is an Ale and create a boolean column for that
summary(as.factor(beer_breweries$Style))

beer_breweries$Style <- as.character(beer_breweries$Style)
beer_breweries$is_ale <- str_detect(beer_breweries$Style, "\\bAle\\b")
summary(beer_breweries$is_ale)
beer_breweries %>%
  filter(is_ale) %>%
  select(Style) %>%
  count(Style)
# We use Regex to find which style is an IPA and create a boolean column for that
beer_breweries$is_IPA <- str_detect(beer_breweries$Style, "\\bIPA\\b")
summary(beer_breweries$is_IPA)
beer_breweries %>%
  filter(is_IPA) %>%
  select(Style) %>%
  count(Style)
  
beer_breweries
#Based on the two previous columns we create a new column which labels the observations as IPA or Ale
for(i in 1:dim(beer_breweries)[1])
{
  if (beer_breweries$is_ale[i]=="TRUE"& beer_breweries$is_IPA[i]== "FALSE")
  {
    beer_breweries$IPA_Ale[i] <- "Ale"
  }
 else if (beer_breweries$is_IPA[i]=="TRUE" & beer_breweries$is_ale[i]=="FALSE")
  {
    beer_breweries$IPA_Ale[i] <- "IPA"
 }
  else
  {
    beer_breweries$IPA_Ale[i] <- "Other"
  }
  
}
beer_breweries
```

#question 8 Pt. 2: Knn for IPA Vs. Ale
```{r}
# We do Knn where we loop for many K  and find the average accuracy with one training and test partion set

#Filter the classifications and then create the training set
beer_breweriesKnn<-beer_breweries %>% filter(IPA_Ale=="IPA"|IPA_Ale=="Ale")
set.seed(6)
splitPerc = .7
trainIndices = sample(1:dim(beer_breweriesKnn)[1],round(splitPerc * dim(beer_breweriesKnn)[1]))
train = beer_breweriesKnn[trainIndices,]
test = beer_breweriesKnn[-trainIndices,]
accs = data.frame(accuracy = numeric(100), k = numeric(100))

#Use Knn for when k = 1-100 and find the accuracy for each value of K
for(i in c(1, 10, 50, 100))
{
styles = knn.cv(beer_breweriesKnn[,c(7,8)],beer_breweriesKnn$IPA_Ale, prob = TRUE, k = i)
  table(beer_breweriesKnn$IPA_Ale,styles)
  CM = confusionMatrix(table(beer_breweriesKnn$IPA_Ale,styles))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
#Plot the accuracy of of value of k
plot(accs$k,accs$accuracy, type = "l", xlab = "k")
accu = accs$accuracy
max(accu)
```


#question 8 Pt. 3: Knn with multiiple iterations for IPA Vs. Ale
```{r}
# We loop for many K where we take the average accuracy of many training and partition sets

#Create 5 different iterations of training and test sets
iterations = 5
numks = 100

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(100), k = numeric(100))
trainIndices = sample(1:dim(beer_breweriesKnn)[1],round(splitPerc * dim(beer_breweriesKnn)[1]))
train = beer_breweriesKnn[trainIndices,]
test = beer_breweriesKnn[-trainIndices,]

# Find the average accuracy with each iteration of the training and tests sets where k = 1-10
for(i in 1:numks)
{
  classifications = knn(train[,c(7,8)],test[,c(7,8)],train$IPA_Ale, prob = TRUE, k = i)
  table(classifications,test$IPA_Ale)
  CM = confusionMatrix(table(classifications,test$IPA_Ale))
  masterAcc[j,i] = CM$overall[1]
}

}

MeanAcc = colMeans(masterAcc)

#Plot the average accuracy of K
plot(seq(1,numks,1),MeanAcc, type = "l",xlab = "k",ylab = "Mean Accuracy")

#Find the maximum average accuracy
max(MeanAcc)
```


#question 9: freeform EDA--most common word used in Beer_Names?

An interesting fact: in beer names, the most common words after the types of beers themselves are colors (red, blonde, brown), then "Wheat", "Porter", and "Summer".

```{r}
#Look at beer names and remove duplicates
head(beer_breweries$Beer_Name)
beer_breweries = beer_breweries %>%
  distinct(Beer_Name)

#put all the names into one big long string of words
Beer_Words = str_c(beer_breweries$Beer_Name, collapse = " ")
Beer_Words

#parse the individual words out into a data frame
Beer_words_distinct = str_match_all(Beer_Words, "(\\b\\w+\\b)")
head(Beer_words_distinct, 15)
Beer_words_df = data.frame(Beer_words_distinct)

#tally the count of each word
Beer_words_df %>%
  count(X1) %>%
  arrange(desc(n))
  
```

